Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-03-26 PAPER.tex
changed 'shows agreement' text to 'shows good agreement'
changed plots to PDFs (except toy model plots)

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/ant_layout_aspect.png}
	\caption{The PAPER-64 antenna layout. We use only the $30$ m East/West baselines for the revised analysis in this 
paper (i.e. the shortest horizontal spacings).}
	\label{fig:ant_layout}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.png}
	\caption{Illustration of the power spectrum amplitude of various power spectrum terms as a function of injected EoR power level summed into the 
data. The net output ($P_{\rm out}=\widehat{\textbf{p}}_r$) is shown in black, the individual terms, as defined by Equation 
\eqref{eq:crossterm},  are shown in blue, red, green; the dotted black line indicates 1:1 input to output mapping.  The 
empirically estimated inverse covariance weighted case used in \citetalias{ali_et_al2015} is on the left and the uniform-weighted 
case is on the right. The details of the simulation used to generate the figure is explained in Section 
\ref{sec:Practice}; here we fit smooth polynomials to our data points to make an illustrative example.   The gray 
horizontal line is the power spectrum value of data alone (doesn't depend on injected power). The green signal-signal 
component is the term used in \citetalias{ali_et_al2015} to estimate signal loss. It is significantly higher 
than the net output because the cross-terms (red) are large and negative (black $=$ green $+$ red $+$ blue) . In the 
regime where cross-correlations between injection and data are not dominant, the cross-terms have a noise-like 
term with width $\sqrt{p_e}\sqrt{p_x}$. However, at power levels comparable to the data (the middle region), the cross-terms can produce large, negative signal due to the couplings between $\textbf{x}$ and $\textbf{e}$ which affect 
$\widehat{\textbf{C}}_{r}$. This causes the difference between the green curve (which exhibits negligible loss at 
the data level) and the black curve (which exhibits $\sim$$3$ orders of magnitude of loss at the data level for 
the weighted case). }
	\label{fig:sigloss_terms}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve_posneg.png}
	\caption{Signal loss transfer functions showing the relationship of $P_{\rm in}$ and $P_{\rm out}$, as defined by Equations \eqref{eq:Pin} and \eqref{eq:sigloss}. Power spectra values (black points) are generated for $20$ realizations of $\textbf{e}$ per signal injection level. Kernel density estimations of the power spectrum transfer functions are shown as colored 
heat-maps on top of the points for the cases of empirically estimated inverse covariance weighted PAPER-64 data (left) and uniform-weighted data (right). The dotted black 
diagonal lines mark a perfect unity mapping, and the solid gray horizontal line denotes the power spectrum value of the data $\widehat{\textbf{p}}_{x}$, from which a posterior distribution for the signal is extracted. From these plots, it is clear that the weighted case results in $\sim4$ orders of magnitude of signal loss at the data level, whereas the uniform-weighted case does 
not exhibit loss. The general shape of these transfer functions are also shown by the black curves in Figure \ref{fig:sigloss_terms} for comparison.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data.png}
	\caption{The power spectrum of PAPER-64 data, weighted by an inverse covariance estimator. The power spectra values pre-signal loss estimation are shown as black and gray points (positive and negative values, respectively, 
with $2\sigma$ error bars). The dashed black curve is the $2\sigma$ upper limit on the EoR signal post-signal loss estimation. The dashed blue line is the uniform-weighted power 
spectrum ($2\sigma$ upper limit). The solid green line is the theoretical $2\sigma$ noise level prediction based on observational 
parameters, whose calculation is detailed in Section \ref{sec:PSSense}.}
	\label{fig:ps2_data}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_modeloop_2panel.png}
	\caption{Power spectra $2\sigma$ upper limits for $k=0.393$ h Mpc$^{-1}$ for fringe-rate filtered PAPER-64 data. Top: Values 
are shown before (gray) and after (black) signal loss estimation as a function of number of eigenmodes of $\widehat{\textbf{C}}$ that 
are down-weighted. This regularization knob is tuned from $0$ modes on the left (i.e. unweighted) to $21$ modes on the right (i.e. the full inverse 
covariance estimator). Over $\sim3$ orders of magnitude of signal loss results when using empirically estimated inverse covariance weighting. Bottom: Power spectrum upper limits before (gray) and after (black) signal loss estimation as a function of identity added to the empirical covariance. This regularization knob is tuned from $0.01\%$  of the $Tr(\widehat{\textbf{C}})$ added to $\widehat{\textbf{C}}$ on the right (i.e. very little regularization) to $100\%$ of $Tr(\widehat{\textbf{C}})$ added to $\widehat{\textbf{C}}$ on the left. Also 
plotted in both panels for comparison are $2\sigma$ power spectrum upper limits for the uniform-weighted case (dashed blue) and inverse variance 
weighted case (red); both are after signal loss estimation. Finally, a theoretical prediction for noise ($2\sigma$ error) is plotted 
as solid green. In the revised PAPER-64 analysis in this paper, we choose to use a regularization scheme of $\widehat{\textbf{C}} \equiv 0.035Tr(\widehat{\textbf{C}}) + \widehat{\textbf{C}}$ as a simple example of regularization that minimizes loss.}
	\label{fig:sigloss_modeloop}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data_add.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_add.png}
	\caption{Power spectrum of PAPER-64 using $\widehat{\textbf{C}}_{eff}$. The weighted power spectrum, pre-signal loss estimation, is shown by the black and gray points (corresponding to positive and 
negative power spectrum values, respectively, with $2\sigma$ error bars). The dashed black curve is the $2\sigma$ upper limit on the EoR signal post-signal loss estimation. There is a small amount of loss using this regularization. The dashed blue line is the uniform-weighted 
power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical $2\sigma$ noise level prediction based on 
observational parameters. This power spectrum result differs from \citetalias{ali_et_al2015} in that it only uses data from one type of baseline ($30$ m 
East/West baselines) instead of three. Major differences from previously published results stem from revisions regarding signal 
loss, bootstrapping, and the theoretical error computation. In Kolopanis et al. (\textit{submitted}), a larger parameter space is explored for regularization and multiple baselines are used in the analysis to produce a best $2\sigma$ power spectrum upper limit of $(240.0$ mK$)^{2}$ at $k=0.3$ $h$Mpc$^{-1}$ and $z=8.37$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim$$10$ in mK.}
	\label{fig:ps1_data}
\end{figure*}

@@

Power spectrum results for the noise simulation, which uses our full power spectrum pipeline, are shown in Figure 
\ref{fig:ps_noise}, where the black and gray points represent pre-signal loss estimated positive and negative power spectrum values, respectively (with 
$2\sigma$ error bars and weighting matrix $\widehat{\textbf{C}}_{eff}$), the dashed black line represents the post-signal loss $2\sigma$ upper limit on the EoR signal, the dashed blue line represents the $2\sigma$ uniform-weighted power 
spectrum limit, and the solid green line denotes our $2\sigma$ theoretical noise prediction as calculated by Equation 
\eqref{eq:sense}. All three are in agreement, validating our analytical thermal noise calculation.

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_noise_add.png}
	\includegraphics[width=0.4\textwidth]{plots/ps2_noise_add.png}
	\caption{Power spectrum estimate for a noise simulation that mimics the noise level of PAPER-64 data. The original weighted 
power spectrum points and their $2\sigma$ errors are show in black and gray (positive and negative values), where we use $
\widehat{\textbf{C}}_{eff}$ to minimize signal loss. The dashed black line is the $2\sigma$ upper limit on EoR post-signal loss estimation. The dashed blue line is the uniform-weighted power spectrum $2\sigma$ upper 
limit. The solid green line is the theoretical $2\sigma$ noise level prediction as calculated by Equation \eqref{eq:sense}. All three 
estimates agree.}
	\label{fig:ps_noise}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.47\textwidth]{plots/null_bls.png}
	\includegraphics[width=0.47\textwidth]{plots/null_lsts.png}
	\includegraphics[width=0.47\textwidth]{plots/null_eo.png}
	\caption{Differenced power spectrum results (with $2\sigma$ errors) for three null tests, where a jackknife is taken along the baseline axis (top left), LST axis (top right), and even/odd Julian date axis (bottom). The results shown are unweighted (no signal loss), where the power spectrum values plotted are computed from the difference between two power spectra produced on either side of the jackknife axis. The gray shaded region in each plot is the estimated $2\sigma$ theoretical noise limit given the parameters of each test. We find that there are no significant systematics that differ between the tests, though there are low level biases at high-$k$ values in both the LST test and even/odd test. All three tests, whether biased or not, have an extra noise bias that is strongly $k$-dependent. Residual foregrounds dominate at $k$-values near the horizon ($k\sim\pm 0.1$ $h$Mpc$^{-1}$), which is not surprising given that foregrounds are brightest in this region post-delay filtering.}
	\label{fig:null}
\end{figure*}

>>>>

\begin{figure}
	\centering
	\includegraphics[trim={0cm 3cm 0cm 3cm},width=\columnwidth]{plots/ant_layout_aspect.pdf}
	\caption{The PAPER-64 antenna layout. We use only the $30$ m East/West baselines for the revised analysis in this 
paper (i.e. the shortest horizontal spacings).}
	\label{fig:ant_layout}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.pdf}
	\caption{Illustration of the power spectrum amplitude of various power spectrum terms as a function of injected EoR power level summed into the 
data. The net output ($P_{\rm out}=\widehat{\textbf{p}}_r$) is shown in black, the individual terms, as defined by Equation 
\eqref{eq:crossterm},  are shown in blue, red, green; the dotted black line indicates 1:1 input to output mapping.  The 
empirically estimated inverse covariance weighted case used in \citetalias{ali_et_al2015} is on the left and the uniform-weighted 
case is on the right. The details of the simulation used to generate the figure is explained in Section 
\ref{sec:Practice}; here we fit smooth polynomials to our data points to make an illustrative example.   The gray 
horizontal line is the power spectrum value of data alone (doesn't depend on injected power). The green signal-signal 
component is the term used in \citetalias{ali_et_al2015} to estimate signal loss. It is significantly higher 
than the net output because the cross-terms (red) are large and negative (black $=$ green $+$ red $+$ blue) . In the 
regime where cross-correlations between injection and data are not dominant, the cross-terms have a noise-like 
term with width $\sqrt{p_e}\sqrt{p_x}$. However, at power levels comparable to the data (the middle region), the cross-terms can produce large, negative signal due to the couplings between $\textbf{x}$ and $\textbf{e}$ which affect 
$\widehat{\textbf{C}}_{r}$. This causes the difference between the green curve (which exhibits negligible loss at 
the data level) and the black curve (which exhibits $\sim$$3$ orders of magnitude of loss at the data level for 
the weighted case). }
	\label{fig:sigloss_terms}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_transfercurve_posneg.pdf}
	\caption{Signal loss transfer functions showing the relationship of $P_{\rm in}$ and $P_{\rm out}$, as defined by Equations \eqref{eq:Pin} and \eqref{eq:sigloss}. Power spectra values (black points) are generated for $20$ realizations of $\textbf{e}$ per signal injection level. Kernel density estimations of the power spectrum transfer functions are shown as colored 
heat-maps on top of the points for the cases of empirically estimated inverse covariance weighted PAPER-64 data (left) and uniform-weighted data (right). The dotted black 
diagonal lines mark a perfect unity mapping, and the solid gray horizontal line denotes the power spectrum value of the data $\widehat{\textbf{p}}_{x}$, from which a posterior distribution for the signal is extracted. From these plots, it is clear that the weighted case results in $\sim4$ orders of magnitude of signal loss at the data level, whereas the uniform-weighted case does 
not exhibit loss. The general shape of these transfer functions are also shown by the black curves in Figure \ref{fig:sigloss_terms} for comparison.}
	\label{fig:sigloss_transfercurve}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data.pdf}
	\caption{The power spectrum of PAPER-64 data, weighted by an inverse covariance estimator. The power spectra values pre-signal loss estimation are shown as black and gray points (positive and negative values, respectively, 
with $2\sigma$ error bars). The dashed black curve is the $2\sigma$ upper limit on the EoR signal post-signal loss estimation. The dashed blue line is the uniform-weighted power 
spectrum ($2\sigma$ upper limit). The solid green line is the theoretical $2\sigma$ noise level prediction based on observational 
parameters, whose calculation is detailed in Section \ref{sec:PSSense}.}
	\label{fig:ps2_data}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_modeloop_2panel.pdf}
	\caption{Power spectra $2\sigma$ upper limits for $k=0.393$ h Mpc$^{-1}$ for fringe-rate filtered PAPER-64 data. Top: Values 
are shown before (gray) and after (black) signal loss estimation as a function of number of eigenmodes of $\widehat{\textbf{C}}$ that 
are down-weighted. This regularization knob is tuned from $0$ modes on the left (i.e. unweighted) to $21$ modes on the right (i.e. the full inverse 
covariance estimator). Over $\sim3$ orders of magnitude of signal loss results when using empirically estimated inverse covariance weighting. Bottom: Power spectrum upper limits before (gray) and after (black) signal loss estimation as a function of identity added to the empirical covariance. This regularization knob is tuned from $0.01\%$  of the $Tr(\widehat{\textbf{C}})$ added to $\widehat{\textbf{C}}$ on the right (i.e. very little regularization) to $100\%$ of $Tr(\widehat{\textbf{C}})$ added to $\widehat{\textbf{C}}$ on the left. Also 
plotted in both panels for comparison are $2\sigma$ power spectrum upper limits for the uniform-weighted case (dashed blue) and inverse variance 
weighted case (red); both are after signal loss estimation. Finally, a theoretical prediction for noise ($2\sigma$ error) is plotted 
as solid green. In the revised PAPER-64 analysis in this paper, we choose to use a regularization scheme of $\widehat{\textbf{C}} \equiv 0.035Tr(\widehat{\textbf{C}}) + \widehat{\textbf{C}}$ as a simple example of regularization that minimizes loss.}
	\label{fig:sigloss_modeloop}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data_add.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_add.pdf}
	\caption{Power spectrum of PAPER-64 using $\widehat{\textbf{C}}_{eff}$. The weighted power spectrum, pre-signal loss estimation, is shown by the black and gray points (corresponding to positive and 
negative power spectrum values, respectively, with $2\sigma$ error bars). The dashed black curve is the $2\sigma$ upper limit on the EoR signal post-signal loss estimation. There is a small amount of loss using this regularization. The dashed blue line is the uniform-weighted 
power spectrum ($2\sigma$ upper limit). The solid green line is the theoretical $2\sigma$ noise level prediction based on 
observational parameters. This power spectrum result differs from \citetalias{ali_et_al2015} in that it only uses data from one type of baseline ($30$ m 
East/West baselines) instead of three. Major differences from previously published results stem from revisions regarding signal 
loss, bootstrapping, and the theoretical error computation. In Kolopanis et al. (\textit{submitted}), a larger parameter space is explored for regularization and multiple baselines are used in the analysis to produce a best $2\sigma$ power spectrum upper limit of $(240.0$ mK$)^{2}$ at $k=0.3$ $h$Mpc$^{-1}$ and $z=8.37$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim$$10$ in mK.}
	\label{fig:ps1_data}
\end{figure*}

@@

Power spectrum results for the noise simulation, which uses our full power spectrum pipeline, are shown in Figure 
\ref{fig:ps_noise}, where the black and gray points represent pre-signal loss estimated positive and negative power spectrum values, respectively (with 
$2\sigma$ error bars and weighting matrix $\widehat{\textbf{C}}_{eff}$), the dashed black line represents the post-signal loss $2\sigma$ upper limit on the EoR signal, the dashed blue line represents the $2\sigma$ uniform-weighted power 
spectrum limit, and the solid green line denotes our $2\sigma$ theoretical noise prediction as calculated by Equation 
\eqref{eq:sense}. All three show good agreement, validating our analytical thermal noise calculation.

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_noise_add.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_noise_add.pdf}
	\caption{Power spectrum estimate for a noise simulation that mimics the noise level of PAPER-64 data. The original weighted 
power spectrum points and their $2\sigma$ errors are show in black and gray (positive and negative values), where we use $
\widehat{\textbf{C}}_{eff}$ to minimize signal loss. The dashed black line is the $2\sigma$ upper limit on EoR post-signal loss estimation. The dashed blue line is the uniform-weighted power spectrum $2\sigma$ upper 
limit. The solid green line is the theoretical $2\sigma$ noise level prediction as calculated by Equation \eqref{eq:sense}. All three 
estimates show good agreement.}
	\label{fig:ps_noise}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.47\textwidth]{plots/null_bls.pdf}
	\includegraphics[width=0.47\textwidth]{plots/null_lsts.pdf}
	\includegraphics[width=0.47\textwidth]{plots/null_eo.pdf}
	\caption{Differenced power spectrum results (with $2\sigma$ errors) for three null tests, where a jackknife is taken along the baseline axis (top left), LST axis (top right), and even/odd Julian date axis (bottom). The results shown are unweighted (no signal loss), where the power spectrum values plotted are computed from the difference between two power spectra produced on either side of the jackknife axis. The gray shaded region in each plot is the estimated $2\sigma$ theoretical noise limit given the parameters of each test. We find that there are no significant systematics that differ between the tests, though there are low level biases at high-$k$ values in both the LST test and even/odd test. All three tests, whether biased or not, have an extra noise bias that is strongly $k$-dependent. Residual foregrounds dominate at $k$-values near the horizon ($k\sim\pm 0.1$ $h$Mpc$^{-1}$), which is not surprising given that foregrounds are brightest in this region post-delay filtering.}
	\label{fig:null}
\end{figure*}
